---
title: Evaluating Methods for Covariate Selection in Experimental Designs
author:
  - name: Kendra Wyant
    corresponding: false
    affiliations:
      - Department of Psychology, University of Wisconsin-Madison
  - name: Lauren Khoury
    corresponding: false
    affiliations:
      - Department of Psychology, University of Wisconsin-Madison
  - name: Markus Brauer
    corresponding: false
    affiliations:
      - Department of Psychology, University of Wisconsin-Madison
  - name: John J. Curtin 
    corresponding: true
    email: jjcurtin@wisc.edu
    affiliations:
      - Department of Psychology, University of Wisconsin-Madison 
abstract: |
  Abstract of paper goes here and can span several lines.
date: last-modified
bibliography: references.bib
bibliographystyle: apa
number-sections: false 
editor_options: 
  chunk_output_type: console
---

<!-- Outlets:
Psychological Science (first choice)
- 2,000 word limit for intro/discussion/notes/acknowledgements/appendices
- No word count for Methods and Results, but they encourage no more than 2,500 words 
- No figure/table limit
- 40 reference limit
- Open access option ($1000 extra)

Plos One
- No word count or figure limit
- Open access journal ($2000 - $3000)

??Psych Methods??
-->
# Introduction

Research in Psychology often uses experimental manipulations to establish causal relationships between a focal independent variable (IV) and a psychological outcome. For example, participants may be randomly assigned to receive a moderate dose of alcohol or placebo to evaluate the effect of alcohol consumption on response to a laboratory stressor (INSERT A COUPLE OF REFS TO OUR STUDIES).  [Markus, can you provide a simple second example that focuses on an IV manipulation that is social psych relevant?)

A valid test of the effect of an IV requires adequate statistical power [@cohenStatisticalPowerAnalysis1988; @cohenPowerPrimer1992]. By definition, low statistical power is associated with a high probability of a Type II error (i.e., failure to reject a false null hypothesis).  Furthermore, significant effects detected by low-powered studies may not replicate (due to low positive predictive value) and estimates of their effect size may be positively biased (due to publication bias).  For all these reasons, methods to increase statistical power are critically important for a robust scientific literature [@cohenStatisticalPowerAbnormalsocial1962; @buttonPowerFailureWhy2013]. 

The use of covariates that are measured at baseline (i.e., prior to random assignment/manipulation of the focal IV) has been shown to increase statistical power and reduce Type II errors by accounting for unexplained variance in the outcome that would otherwise be treated as noise in the analysis [@thompsonCovariateAdjustmentHad2015; @hernandezCovariateAdjustmentRandomized2004; @hernandezRandomizedControlledTrials2006; @kahanRisksRewardsCovariate2014; @vanbreukelenANCOVAChangeBaseline2006; @egbewaleBiasPrecisionStatistical2014]. For example, in the previously described test of the effect of alcohol on stress response, baseline measures of both trait anxiety and fearfulness could be included as covariates.  These measures would not be systematically correlated with the IV because they are collected prior to random assignment.  They would also be expected to correlate with laboratory stress response based on prior research. Given this, the addition of these measures as covariates in the analysis could reduce variance in stress response that is unrelated to the manipulation to increase statistical power to test the IV effect.

However, selection of the optimal covariates to include to increase statistical power is not straightforward. Historically, researchers often iterated over a large set of available covariates and included only those covariates that lowered the *p*-value for the test of their focal IV.  Today, we recognize that this approach, known as *p*-hacking [@simonsohnPcurveKeyFiledrawer2014], seriously undermines the statistical validity of the analyses by substantially increasing the probability of Type I errors (i.e., rejecting the null hypothesis when it is not false [@simmonsFalsepositivePsychologyUndisclosed2011; @ioannidisWhyMostPublished2005; @schelchterPosthocSelectionCovariates1985; @beachChoosingCovariatesAnalysis1989]). Statistically valid analyses require both low probability of Type II errors and control of Type I errors at the specified value of alpha (e.g., 0.05).  

Given that *p*-hacking compromises statistical validity, we are left with questions about appropriate covariate selection approaches that both minimize Type II errors and control Type I errors at alpha.  A priori, the researcher could decide to avoid the use of covariates altogether with no risk to Type I error control.  But statistical power would likely be lost by this decision.  At the other extreme, the researcher could plan to include all available covariates that are reasonably believed to account for unexplained variance in the outcome in the statistical analysis. This decision, if made a priori, would also not be expected to inflate the Type I error rate. However, this approach may not yield optimal statistical power because a subset of these covariates may not manifest strong, or any, relationship with the outcome in any particular study if they have not been previously well-established as robust predictors.  Furthermore, available covariates are often correlated with each other (e.g., trait anxiety and fearfulness from the previous example), which reduces their potential benefit in reducing noise relative to their cost to degrees of freedom in the analysis.  

Given these possibly sub-optimal previous options, it may be that a data-driven approach to covariate selection could improve statistical validity relative to the a priori all or no covariate approaches.  An optimal data-driven covariate selection method may be able to select the covariates that provide the greatest increase in power (e.g., covariates highly correlated with the outcome [@kahanRisksRewardsCovariate2014]) with the lowest cost to degrees of freedom (e.g., by including only covariates that contribute to unique reductions in variance). Of course, statistically valid data-driven approaches must also not meaningfully inflate the Type I error rate, as occurs with *p*-hacking. 

In this study, we use simulation methods to evaluate six candidate data-driven approaches to select an optimal set of covariates when testing the effect of a focal manipulated IV.  For each approach, we use covariate "screening models" to identify which covariates display significant relationships with the outcome variable and will be included in the subsequent primary analysis that tests the focal IV. Across these screening models, we consider three modeling approaches (i.e., linear models that screen each covariate separately in its own model, linear models that screen all covariates simultaneously in a single model, and LASSO [@tibshiraniRegressionShrinkageSelection1996a] models that screen all covariates simultaneously in a single model).  Across these modeling approaches, we also either include or exclude the focal IV in the screening models for a total of six candidate data-driven approaches.  We benchmark these candidate data-driven approaches against a priori "all covariates" and "no covariates" approaches as well as the *p*-hacking approach (a statistically invalid data-driven approach).  For all approaches we quantify the probability of Type I and Type II errors across a variety of research contexts that vary by sample size, number of available covariates, strength of covariate effects, effect size for the IV and other relevant context characteristics.  We use results from these simulations to provide clear and accessible guidance about how best to select an optimal set of covariates to increase statistical power when testing the effect of a manipulated IV in various research contexts.


# Methods

## Covariate Selection Approaches

As described above, we evaluated six candidate data-driven and three benchmark (*p*-hacking, all covariates, no covariates) approaches to select covariates to use in a primary analysis that tested the effect of a focal dichotomous IV (notated as $X$).  A summary of these nine approaches is presented in @tbl-methods.

For the six candidate data-driven approaches, we selected covariates for the primary analysis of $X$ by fitting preliminary covariate screening models that used either the linear model (i.e., ordinary least squares regression) or LASSO. Across these approaches we considered screening models that either included or did not include $X$ in the screening model.  

For the candidate approaches that used general linear models for screening, we considered approaches where all covariates were screened simultaneously in one model or where the covariates were screened one at a time in separate models. Across these approaches, a covariate was selected to be included in the primary analysis if the *p*-value for that covariate was less than 0.05 in the screening model that evaluated it. The *p*-value for $X$ was not used and $X$ was only included to parse its variance from $Y$ when evaluating the covariates.  

For the candidate approaches that used LASSO (with and without $X$) for screening, the screening model simultaneously included all covariates. Also, for the LASSO screening model that included $X$, the regularization penalty (lambda) for $X$ was set to 0 so that $X$ was always retained in the screening model.  Covariates were selected to be included in the primary analysis if they were retained by LASSO in the screening model (i.e., did not have their parameter estimate set to 0).

We benchmarked these six candidate data-driven approaches against a statistically invalid data-driven approach (*p*-hacking) and two a priori selection approaches that did not use the data to screen and select covariates.  The *p*-hacking approach used separate screening models for each covariate and each of these models also included $X$.  Covariates were selected to be included in the primary analysis if the *p*-value for $X$ was reduced (relative to its bivariate relationship with $Y$) in the screening model that included that covariate. The all covariates approach and the no covariates approach did not use screening models but instead used either all or none of the available covariates in the specific research context in the primary analysis of $X$. 

 
{{< embed notebooks/mak_tables.qmd#tbl-methods >}}


## Research Contexts
We manipulated several characteristics designed to mimic a range of research contexts that researchers in psychology might encounter in their studies. We crossed all levels of each characteristic to create a total of 540 unique research contexts to allow results from our simulations to generalize across the diversity of contexts within which psychology studies are situated. A summary of these characteristics is presented in @tbl-dictionary. They include:

1. **The population parameter for the focal IV ($\beta_X$)**. We selected values that represent null ($\beta_{X} = 0$), small ($\beta_{X} = 0.2$) and medium ($\beta_{X} = 0.5$) effect sizes for the IV. Given that the standard deviation of the outcome variable was set to 1 in all simulations, these parameter values correspond to null, small, and medium Cohen's d effect sizes [@cohenStatisticalPowerAnalysis1988].

2. **The sample size ($N$)**. We chose values for the number of observations that pertain to common sample sizes in experimental research in psychology: 50, 100, 150, 200, 300, and 400 observations.

3. **The number of covariates available ($N_{cov}$)**. We selected a wide range of possible scenarios: 4, 8, 12, 16, or 20 available covariates.

4. **The percentage of covariates with non-zero correlations with the outcome ($\%_{cov}$)**. We simulated research contexts where 25%, 50% or 75% of the available covariates had non-zero correlations with the outcome in the population.  This allowed us to model situations where the researcher has an exploratory vs. more well-established set of covariates at their disposal. 

5. **The magnitude of the correlation between the covariates and $Y$ ($r_{cov-Y}$)**. All non-zero covariates (see characteristic 4) were given a moderate or large relationship to the outcome $Y$ (i.e., $r_{cov-Y}$ = 0.3 or 0.5, respectively). Given that these covariates are all correlated with the outcome, it is expected that they would typically also be correlated with each other. Therefore, we assigned a moderate 0.3 correlation among each of these non-zero covariates. 


{{< embed notebooks/mak_tables.qmd#tbl-dictionary >}}


## Data Analytic Plan

All data analyses were done in R using the Tidyverse ecosystem [@kuhnTidymodelsCollectionPackages2020] and the RStudio IDE [@rstudioteamRStudioIntegratedDevelopment2020]. Simulations were run using high-throughput computing resources provided by the University of Wisconsin-Madison Center for High Throughput Computing [@chtc]. 

We simulated 40,000 datasets within each research context. For each simulated dataset, we generated scores for a dichotomous focal independent variable ($X$), varying numbers of quantitative covariates ($C$), and a quantitative outcome ($Y$).  To generate these datasets, we first created $N$ scores for a set of multivariate normal distributed covariates and $Y$, with their variances = 1 and the number of covariates, and the covariances among the covariates and with $Y$ dictated with the research context (see points 3-5 above).  Consistent with the manipulation of the IV following measurement of the baseline covariates, the effect of $X$ was ($\beta_X * X$) was then added into $Y$.

Below, we display an exemplar covariance matrix and subsequent R code to generate a single simulated dataset for a research context with a small population effect size for the IV ($\beta_{X} = 0.2$), a sample size of 100, and four covariates, where 50% of the covariates have a 0.50 correlation with the outcome and 0.30 correlations with each other.

This is the relevant covariance matrix (sigma).
```{r}
#| echo: false

sigma <- matrix(c(
  1.0, 0.5, 0.5, 0.0, 0.0,
  0.5, 1.0, 0.3, 0.0, 0.0,
  0.5, 0.3, 1.0, 0.0, 0.0,
  0.0, 0.0, 0.0, 1.0, 0.0,
  0.0, 0.0, 0.0, 0.0, 1.0
), nrow = 5, byrow = TRUE)

rownames(sigma) <- c("Y", "C1", "C2", "C3", "C4")
colnames(sigma) <- c("Y", "C1", "C2", "C3", "C4")


print(formatC(sigma, format = "f", digits = 1), quote = FALSE)
```

The covariance matrix can then be used to generate a multivariate normal dataset of $Y$ and $C_1...C_4$ for 100 observations where all variables have a mean of 0.
Following this, the effect of the IV ($X$) can be added into $Y$.

```{r}
#| echo: true

df <- MASS::mvrnorm(n = 100, mu = c(0, 0, 0, 0, 0), Sigma = sigma) # <1>

x <- rep(c(-.5, .5), each = 50) # <2>
df[, 1] <- df[, 1]+ .2 * x # <3>
```

1. Generate $Y$ and all covariates using a multivariate normal distribution with all means of 0 and variances/covariances from sigma displayed above.
2. Create a dichotomous focal independent variable ($X$). 
3. Add in $X$ effect on $Y$.

For each simulated dataset in each research context, we used each of the 9 covariate selection approaches to determine which covariates to use in the primary analysis to test the effect of $X$ on $Y$. For example, if the selection approach selected covariates 1 and 4, the primary analysis would be $lm(Y \sim X + C_1 + C_4, \text{data} = \text{df})$.  From each primary analysis, we saved the research context characteristics, the covariate selection approach used, and the parameter estimate for $X$, *p*-value for its statistical test against 0, and the conclusion (reject or fail to reject) using alpha = 0.05.  

For research contexts where the population parameter for $X$ was 0 (i.e., $X$ has no effect on $Y$), we assessed and reported the Type I error rates associated with each covariate selection approach by research context characteristics. For research contexts where the population parameter for $X$ was 0.2 or 0.5 (i.e., $X$ has an effect on $Y$), we assessed and reported the Type II error rate for each covariate selection method by research context characteristics.  We also provided descriptive statistics and sampling distributions (from the 40,000 simulations) of the parameter estimate for $X$ across research contexts for each covariate selection approach, separately by true effect size for $X$ (0, 0.2, and 0.5). An interactive shiny dashboard of mean error rates for all combinations of research context variables is accessible through our OSF page ([https://osf.io/zh5gn/](https://osf.io/zh5gn/)).

# Results

## Type I Error

Given our use of simulated datasets, we were able to set the true population parameter for $X$ to be zero (i.e., $\beta_X = 0$). Therefore, any significant result was a Type I error. @fig-bar-1 shows the average Type I error rate across all research contexts for each approach. The *p*-hacking approach for selecting covariates, unsurprisingly led to a highly inflated mean Type I error rate across contexts, consistent with the extant research published in the last several years on the connection between researcher degrees of freedom and false positives<!--kendra Insert refs-->}. The other approaches were at or near the expected error rate of alpha = .05 across the relatively broad range of research context characteristics we considered. Specifically,  the two a priori selection approaches and the data-driven selection approaches that did not include $X$ in the screening model displayed mean Type I error rates of exactly .05.  The data-driven approaches that did include $X$ in the screening model displayed slightly increased (<= .06) mean Type I error rates across contexts.  


{{< embed notebooks/mak_figures.qmd#fig-bar-1 >}}


We next assessed how Type I error rates varied by specific dimensions of our research contexts and noted several patterns (@fig-type1-panel). First, the magnitude of Type I error rate inflation for the *p*-hacking approach was substantially increased with increasing number of available covariates, larger correlations between the covariates and $Y$ (i.e., $r_{cov-Y}$ = 0.5), higher proportion of available covariates with non-zero correlations with $Y$, and smaller sample sizes.  The Type I error rate for *p*-hacking reached 0.462 in the worst case context (N = 50, 20 available covariates with $r_{cov-Y}$ = 0.5 for 75% of the covariate set; see Shiny dashboard for other contexts).  

In contrast, to the *p*-hacking approach, the Type I error rates remained constant at 0.05 across all research context dimensions for the a priori approaches and the data-driven approaches that did not include $X$.  Among the data-driven approaches that did include $X$, small increases in Type I error rates were observed as the sample size decreased and the number of available covariates increased.  However, Type I error rate inflation remained modest for these approaches even at the extremes of these two dimensions (up to 0.074 at N = 50 and up to 0.066 with 20 covariates in the simultaneous linear model approach). And in the worst case context at the intersection of all dimensions, error rates for these screening approaches that included $X$ only reached 0.072 for separate linear models, 0.096 with a simultaneous linear model and 0.089 with a simultaneous LASSO.  Neither the magnitude $r_{cov-Y}$ nor the proportion of available covariates with non-zero correlations with $Y$ appears to have any systematic impact on the Type I error rates for the data-driven approaches that included $X$. 


<!--Kendra, figure axis and title needs to be updated to use percent rather than proportion of good covariates as that is how we describe that dimension in the text-->
{{< embed notebooks/mak_figures.qmd#fig-type1-panel >}}


## Type II Error

We simulated datasets with two possible non-zero population parameters for the effect of $X$ on $Y$ ($\beta_X = 0.2$, $\beta_X = 0.5$). Given that we set the standard deviation of $Y$ to equal 1, these represented contexts where $X$ had a small (Cohen's d = 0.2) or moderate (Cohen's d = 0.5) effect on $Y$.  In these simulations, non-significant effects for $X$ were, by definition, Type II errors. Given the substantial problems with Type I error rate inflation noted above, we do not consider the *p*-hacking approach to be viable and therefore did not evaluate Type II error rates for this approach.  

@fig-bar-2 shows the mean Type II error rate across all research contexts for each approach. The a priori no covariates approach displayed the highest average Type II error (0.449), which reinforces the value of using covariates to increase statistical power to detect true effects.  Among the approaches that used covariates, the candidate data-driven approaches generally produced modestly lower average Type II error rates than the a priori all covariates approach.  Furthermore, the data-driven approaches that included $X$ displayed modestly lower mean Type II error rates than the comparable approaches that did not include $X$ in the screening model.  


{{< embed notebooks/mak_figures.qmd#fig-bar-2 >}}

NEED TO UODATE THIS NEXT PARAGRAPH

We next assessed how Type II error rates varied by specific dimensions of our research contexts and noted several patterns (@fig-panel-2). First, it is clear that the no covariates approach was always associated with the highest Type II error rates across all values of all dimensions of the research contexts. Second, the all-covariates approach tended to under-perform (i.e., higher Type II error rates) relative to the candidate data-driven approaches when the sample size was low, the number of available covariates was high, or the percentage of good covariates was low.  Similarly, the data-driven approaches that did not include $X$ tended to under-perform relative to the matched approaches that did include $X$ in these same contexts (low N, high number of covariates, low percentage of good covariates).

<!--Kendra, figure axis and title needs to be updated to use percent rather than proportion of good covariates as that is how we describe that dimension in the text-->

<!--Kendra, can we consider truncating the y axis too?  Starting at a value great than 0 would allow it to be easier to tell the lines apart.  Can we see those graphs?-->
{{< embed notebooks/mak_figures.qmd#fig-panel-2 >}}


## Parameter Estimates

When the assumptions of the linear model are met , parameter estimates are unbiased (i.e. mean of the parameter estimate's sampling distribution equals the value of the population parameter) and their sampling distributions follow the t-distribution.  @tbl-est shows the mean parameter estimate for $\beta_X$ from our primary analysis for each covariate screening approach across research contexts (with 40,000 sample parameter estimates per context), separately by population effect size ($\beta_X = 0$, $\beta_X = 0.2$, $\beta_X = 0.5$). 


{{< embed notebooks/mak_tables.qmd#tbl-est >}}


We also display the associated sampling distributions for these parameter estimates for each approach and population effect size in @fig-panel-2.  


{{< embed notebooks/mak_figures.qmd#fig-distribution-bx >}}

EDIT HERE

<!--need to consider what to report here.  Should we only report the patterns or also the explanations for why?  Explanations could be moved to discussion but that would cost us words there!-->
All approaches yielded unbiased estimates when the true population effect size was 0.  However, the shape of the sampling distribution for the *p*-hacking approach was clearly distorted (bi-modal). This distortion results from this method selecting covariates that magnify the effect of $X$ in any specific sample.  When sampling error produces a negative $b_X$ in a sample, covariates that make that parameter estimate more negative will be selected, and when sampling error produces a positive $b_X$ in a sample, covariates that make that parameter estimate more positive will be selected.  This is what produces the intolerably high Type I error rates associated with that approach.

HERE

When there is a true non-zero population effect for $X$, *p*-hacking yielded substantial optimistic (over-estimation) bias in the parameter estimates.  In this instance, the sample $b_X$ was typically positive to start (given the positive population parameter) and covariates were selected only if they made that parameter estimate even more positive.

Of course, the a priori approaches produced unbiased parameter estimates because the linear model used the the primary analysis of $X$ was In contrast, when there is a non-zero effect.  the a priori approaches and the candidate data-driven approaches that include $X$ yielded unbiased parameter estimates, whereas there was slight conservative (under-estimation) bias in the parameter estimates for the data-driven approaches that do not include $X$.  In contrast, 


unpack bias for models without X



# Discussion

Our findings reinforce the cautionary warning by Simmons and colleagues [@simmonsFalsepositivePsychologyUndisclosed2011]: *p*-hacking severely inflates Type I error rate. Simmons et al. simulated an average Type I error rate of 11.7% for a treatment effect when selecting between two potential covariates (*N*=40). In our closest research context (*N*=50, 4 covariates), we observed similar rates of 10.3% (range: 5.6%-15.2%)<!--KW: these ranges are across simulations. Maybe we don't need them here? Just average?-->. However, our simulations suggest this estimate may gravely underestimate the magnitude of the problem of *p*-hacking. It is more realistic that researchers can choose from several to many available candidate covariates. Even when the number of available covariates is modest (i.e., 8-12), our results show that Type I error rates increase to 19.1% (range: 12.1%-27.5%) across contexts. When selecting among 16-20 covariates, Type I error rates, on average, reach nearly a 3-fold increase from our 4-covariate context (29.4%, range: 21.7%-38.3%). 

When covariates were not used at all (i.e., a priori no covariates method), simulated models yielded an average Type I error rate of .05 and an unbiased parameter estimate of the $X$ effect. However, this approach had a meaningfully higher Type II error rate compared to the other methods. This is not surprising. Adding covariates correlated with the outcome and uncorrelated with $X$ will account for additional variance in the outcome, thus increasing power (or lowering Type II error rate). Perhaps then it is also not surprising that the research contexts with highest discrepancies in Type II error rates between the no covariates method and selection methods include situations with moderate to large sample sizes (*N*=150-400) where degrees of freedom can afford the cost of  additional parameters, and in situations where there are many covariates to choose from with strong correlations to the outcome, where covariates have the most opportunity to reduce unexplained variance in the outcome.  We believe that these data provide strong evidence to support the use of covariates when available.

In many instances, using all available covariates selected a priori to data analysis is a reasonable option. Our simulated models showed consistent average Type I error rates of .05 and unbiased parameter for $X$. Additionally, on average Type II error rates were only about 1.5-2% higher than some of the best performing data-driven selection methods (i.e., all covariates LASSO with $X$ and single covariate linear model with $X$). Using all available covariates is easiest because it eliminates the need for selection. However, in certain contexts, such as with small sample sizes or large numbers of covariates this difference is no longer trivial, with some data-driven selection methods yielding Type II error rates as much as 7% lower than the all covariates method. These findings are concerning in that it is not uncommon in psychological research to have 20+ covariates available. Moreover, the high cost of running experimental studies may confine researchers to smaller than ideal sample sizes. Moreover, even with an adequate sample size and modest number of covariates, it may not be wise to walk away from any power to detect an effect. 
<!--JC note: Anchor Type II benefit?  small but how to think about it?   N increase but not linear so specific N?-->

Among the three data-driven selection methods our simulations showed that when the methods did not include the focal variable, $X$, in the model they produced biased parameter estimates. This bias results from the models favoring selecting covariates that are correlated with $X$ in the sample despite being uncorrelated in the population. In other words, the models are accounting for variance in the outcome related to $X$ by capitalizing on spurious correlations. This bias was largest for the all covariates linear model and trivial with the single covariate linear model.

However, data-driven selection methods that included $X$ in the model produced larger Type I error inflation. 
<!--JC notes: 
- Anchor Type I inflation - need simulation
- Why? Flexibility of models? Seen across all methods that include X.-->

Our findings suggest the all covariates linear model is not an optimal data-driven covariate selection method. On average it performed worse than the other data-driven methods. When no true effect existed the all covariates linear model with $X$ yielded the highest Type II error compared to the other data-driven selection methods. When there was a true effect, the all covariates linear model with and without $X$ showed higher Type II error rates compared to the respective single covariate linear and all covariates LASSO models, with and without $X$. Finally, the all covariates model with $X$ produced the most biased parameter estimates, compared to the other data-driven selection methods, when true effects existed. 

The absolute best method for Type II error is all covariates LASSO with $X$ included in the model. LASSO is a machine learning algorithm that may be unfamiliar to some psychology researchers. As such, this data-driven selection method may feel overly complicated. We provide code in the supplement<!--need to add this if JC agrees--> to demonstrate how LASSO can be easily implemented in R for covariate selection.

However, the simple single covariate linear model with $X$ included in the model may be an equally viable alternative to LASSO. This method performed almost as good as LASSO with regard to Type II error and had lower Type I error compared to LASSO. Moreover, this method is easy to implement as it can be run as a table of correlation or partial correlation values. Both of these methods yielded unbiased parameter estimates and are likely comparable.  Researchers can weigh the trade off between Type I and Type II error and the complexity of these methods in choosing which data-driven selection method to use. And of course, this decision should be pre-registered to avoid introducing additional researcher degrees of freedom into the analyses. 

This study demonstrated the gains in power when using covariates measured before assignment. These findings are likely still applicable to covariates uncorrelated to the focal variable and empirically related to the outcome but measured after assignment. Another use of covariates that we did not address in this study is to statistically control for confounding effects by including additional variables related to both the focal variable and the outcome. This is common practice in situations where $X$ cannot be manipulated (e.g., controlling for gender when looking at the effects of parenting style on child school performance). This use of covariates, however, is complicated and covariates must be selected carefully based on theory and causal structure. Unfortunately, controlling for a confounding variable is also not as strong as people tend to believe. Controlling for a confounding variable with a covariate does not fully eliminate its influence but removes the partial effect on the outcome measurable by the model. Given the complex nature of these types of covariates, we considered them to fall outside the scope of the current paper and we plan to address them in a future simulation paper.


# References
