---
title: "Analysis of Covariate Results"
author: "Lauren Khoury"
date: "`r lubridate::today()`"
format: 
  html: 
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

Set up.

```{r set up}

# load packages + set ggplot theme

library(dplyr) |> suppressMessages()
library(skimr)
library(purrr)
library(ggplot2)

theme_set(theme_classic())

# format path 

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true") |> suppressMessages()

rdrive_path <- format_path("cov/raw_data")

```

Read in data.

```{r read in data}

data <- read.csv(here::here(rdrive_path, "batch_results.csv")) |> 
  mutate(model = gsub("[ -]", "_", model)) |> 
  glimpse()

```

Look at proportion significaint ($p < 0.05$) by model.

```{r p val by model}

data |> 
  group_by(model) |> 
  summarise(prop_sig = mean(p_value < 0.05))

```

Look at proportion significaint ($p < 0.05$) by model and research setting.

```{r p val res setting}

data |> 
  group_by(model, n_obs, b_x, n_covs, b_cov, p_good_covs, r_cov) |> 
  summarise(prop_sig = mean(p_value < 0.05),
            .groups = "drop") |> 
  View()

```

Function to plot combined histogram and density plot.

```{r function plot hist dens}

plot_hist_density <- function(data, model_name) {
  ggplot(data |> filter(model == model_name), aes(x = estimate)) +
    geom_histogram(aes(y = after_stat(density)), bins = 10, fill = "grey", color = "black") +
    geom_density(fill = "blue", alpha = 0.25, lwd = 1) + 
    labs(title = paste("Density Plot:", model_name), 
         x = "Beta_X Estimate", y = "Density")
}

```

Plotting for all 5 models.

```{r plot hist dens}

plot_hist_density(data, "lm_no_covs")
plot_hist_density(data, "lm_all_covs")
plot_hist_density(data, "lm_p_hacked")
plot_hist_density(data, "lm_partial_r")
plot_hist_density(data, "lm_lasso")

```
