---
title: "Separate LM with X"
format: html
editor_options: 
  chunk_output_type: console
---


**This script demonstrates how to use separate linear models with X as covariate screening models for selecting covariates.**


### 1. Read in simulated data set. 

```{r}
data <- readr::read_csv("./simulated_data.csv", show_col_types = FALSE) 
```

Get number of covariates in data.
```{r}
n_covs <- names(data) |> 
  stringr::str_subset("^c") |> 
  length()
```



### 2. Select covariates.

These screening models select covariates that are significant on y, controlling for x. They consider each covariate one at at a time and includes it in the final model only if it is significant.

```{r}
# empty vector of covariates that are significant on y, controlling x
covs_added <- character(0) 

# loop through all covariates and test significance when controlling for X
for(i in 1:n_covs) {
  ci <- grep("^c", names(data), value = TRUE)[i]
  formula_1cov <- reformulate(termlabels = c("x", ci), response = "y")
  lm_1cov <- lm(formula = formula_1cov, data = data)
  p_1cov <- lm_1cov |> 
    broom::tidy() |> 
    dplyr::filter(term == ci) |> 
    dplyr::pull(p.value)
    
  if(p_1cov < .05) {
    covs_added <- c(covs_added, ci)
  }
}

covs_added
```



### 3. Fit final linear model with covariates selected by separate lm with X. 

```{r}
formula_final <- reformulate(termlabels = c("x", covs_added), response = "y")

lm(formula = formula_final, data = data)
```

