---
title: "Simultaneous LASSO with X"
format: html
editor_options: 
  chunk_output_type: console
---


**This script demonstrates how to use a simultaneous lasso with X covariate screening model for selecting covariates.**


### 1. Read in simulated data set. 

```{r}
data <- readr::read_csv("./simulated_data.csv", show_col_types = FALSE) 
```

Get number of covariates in data.
```{r}
  n_covs <- names(data) |> 
    stringr::str_subset("^c") |> 
    length()
```



### 2. Tune LASSO model.

Tune (across 100 bootstraps) and fit a best LASSO model using x and all covariates with no penalty applied to x (so it is never dropped).
```{r}
# create bootstrap splits
set.seed(102030)
splits_boot <- data |> rsample::bootstraps(times = 100)
  
# create hyperparameter tuning grid for penalty/lambda values
grid_penalty <- tidyr::expand_grid(penalty = exp(seq(-8, 8, length.out = 1000)))
  
# Fit LASSO models. No need to standardize covariates because all have same variance and penalty will not be applied to x
fits_lasso <- parsnip::linear_reg(penalty = tune(), mixture = 1) |> 
  parsnip::set_engine("glmnet",
                      penalty.factor = c(0, rep(1, n_covs))) |>
  tune::tune_grid(preprocessor = recipes::recipe(y ~ ., data = data),
                  resamples = splits_boot, 
                  grid = grid_penalty, 
                  metrics = yardstick::metric_set(yardstick::rmse))
```

### 3. Fit best LASSO model on all data.

```{r}
fit_best_lasso <-
  parsnip::linear_reg(penalty = tune::select_best(fits_lasso, metric = "rmse")$penalty, 
                      mixture = 1) |> 
  parsnip::set_engine("glmnet",
                      penalty.factor = c(0, rep(1, n_covs))) |> 
  parsnip::fit(y ~ ., data = data)
```



### 4. Select covariates that had non-zero coefficients in the best LASSO model.

```{r}
covs_added <- fit_best_lasso |> 
  broom::tidy() |> 
  dplyr::filter(estimate != 0) |> 
  dplyr::filter(stringr::str_starts(term, "c")) |> 
  dplyr::pull(term)

covs_added
```


### 5. Fit linear model with covariates selected by LASSO. 

```{r}
formula_final <- reformulate(termlabels = c("x", covs_added), response = "y")
  
lm(formula = formula_final, data = data)
```

